\documentclass{article}
\usepackage{amsmath}
\usepackage{collcell,booktabs,siunitx}
\usepackage{xparse}

\title{E coli by numbers}
\author{Donghoon Lee}
\data{\today}

\begin{document}
    \maketitle

    section{General}
    \begin{itemize}
        \gdef\NA{#6.022*10^{23}}
        \item Avogadro's constant = 6.022 x 10^{23} ≈ 6 x 10^{23}
        \item total protein density ≈ 2~4 x 10^{6} proteins / μm^{3}
        \item mean diameter of protein ≈ 4~5 nm
        \item 1 base pair volume ≈ 1 nm^{3}
        \item time scale for transitions between protein states (active/inactive) ≈ 1~100 μs
        \item time scale for transcription factor binding to DNA site ≈ 1 s
        \item mutation rate ≈ 10^{-8} ~ 10^{-10} / bp / replication
    \end{itemize}

    section{Ecoli}
    \begin{itemize}
        subsection{Physical Specification}
        \gdef\Volume{#1*10^{-18}}
        \item cell volume ≈ 0.3~3 μm3    
        \item cell dimension = 2 μm x 1 μm x 1 μm, rod shape
        \item total weight ≈ 9.5?10^-13g % Neidhardt_1996
        \item total dry weight ≈ 2.8 x 10^{-13} g (30\% of total mass, where 70\% of it being water) % Neidhardt_1996
        
        subsection{Chemical Composition}
        \item water ≈ 2 x 10^{10}
        \item inorganic ion ≈ 1 x 10^{8} ≈ 4\% of dry weight % Frederick_1996
        \item lipids ≈ 5 x 10^{7} ≈ 10\% of dry weight % Frederick_1996
        \item DNA / cell ≈ 4.6 x 10^{6} bp ≈ 3\% of dry weight % Frederick_1996
        \item RNA / cell ≈ 2 x 10^{3} ≈ 20\% of dry weight % double check! Frederick_1996
        \item proteins / cell ≈ 3 x 10^{6} ≈ 55\% of dry weight % Frederick_1996
        
        \item RNA polymerases / cell ≈ 2 x 10^{3} % BremerDennis_1996
        \item ribosomes / cell ≈ 1 x 10^{4}

        subsection{Genetic Info}
        \item genome size ≈ 4.6 Mbp
        \item # of protein coding genes ≈ 4300
        \item regulator binding site length ≈ 10~20 bp
        \item promoter length ≈ 100 bp
        \item gene length ≈ 1000 bp

        subsection{Dynamics}
        \item transcription rate < 1 min (80 nts/s)
        \item translation rate < 1 min (20 aa/s)
        \item mRNA life time ≈ 3 min
        \item protein life time ≈ 1 hr
        \item minimal doubling time ≈ 20 min

        subsection{Molecular Kinetics}
        \item diffusion time of protein across cell ≈ 0.01 s
        \item diffusion time of small molecule across cell ≈ 0.001 s
        \item time scale for equilibrium binding of small molecule to protein ≈ 1 ~ 1000 ms (1uM ~ 1nM affinity)

    \end{itemize}    

    section{Ecoli Metabolite Concentration}

    \begin{tabular}{ ||l|l|l| } % |l|l|l||l|l|l||} 
        \hline
        \multicolumn{3}{|l|}{Glucose Metabolism} \\
        \hline
        Metabolite & mol L^{-1} (= M) & # of molecules \\ 
        \hline
        ATP     & 9.6 x 10^{-3} & \STcopy{v}{round (b3 * \NA * \Volume)}    \\ 
        ADP     & 5.6 x 10^{-4} &                                           \\ 
        NAD+    & 2.6 x 10^{-3} &                                           \\ 
        \hline

        % check build before adding more

    \end{tabular}

    section{Reference}
    \begin{itemize}
        \item Milo, R., & Phillips, R. (2015). Cell biology by the numbers. CRC Press.
        \item Bennett, B., Kimball, E., Gao, M. et al. Absolute metabolite concentrations and implied enzyme active site occupancy in Escherichia coli. Nat Chem Biol 5, 593–599 (2009). https://doi.org/10.1038/nchembio.186
        \item Bremer, H., Dennis, P. P. (1996) Modulation of chemical composition and other parameters of the cell by growth rate. Neidhardt, et al. eds. Escherichia coli and Salmonella typhimurium: Cellular and Molecular Biology, 2nd ed. chapter 97, p.1559, Table 3 and note 'b' beneath table
        \item Frederick C. Neidhardt AND H. Edwin Umbarger, Chemical Composition of Escherichia coli, chapter 3 in Neidhardt F.C. Escherichia coli and Salmonella: Cellular and Molecular Biology. 2nd edition. Vol 1. American Society of Microbiology (ASM) Press 1996. p.2 table 1
    \end{itemize} 

\end{document}

